Modelling state-switching functional data with hidden Markov models: supplement A (case study)

### 2021-07-07 Report/Supp_Case.tex
Update Simulation Study in Report
Update Report

    \section{Likelihood Surfaces - CarHHMM-DFT}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the dive duration parameters $\mu_Y^{(i)}$ and $\sigma_Y^{(i)}$ for dive types $i = 1,2$. All other parameters and probability transition matrices are assumed to be the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is added to the negative log-likelihood of the maximum likelihood estimate $(\hat \mu_Y,\hat \sigma_Y)$, which is denoted by a red star.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the coarse-scale probability transition matrix parameters $\eta_{1,2}$ and $\eta_{2,1}$. Recall that $\Gamma_{i,j} = \exp(\eta_{i,j}) / \sum_k \exp(\eta_{i,k})$. All other parameters and probability transition matrices are assumed to be the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is added to the negative log-likelihood of the maximum likelihood estimate $(\hat \eta_{1,2},\hat \eta_{2,1})$, which is denoted by a red star.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the wiggliness parameters $\mu_W^{*(\cdot,i^*)}$ and $\sigma_W^{*(\cdot,i^*)}$ for subdive types $i^* = 1,2,3$. All other parameters and probability transition matrices are assumed to be the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is added to the negative log-likelihood of the maximum likelihood estimate $(\hat \mu_W^{*(\cdot,i^*)}, \hat \sigma_W^{*(\cdot,i^*)})$, which is denoted by a red star.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the $x$-acceleration parameters $\mu_{A_x}^{*(\cdot,i^*)}, \sigma_{A_x}^{*(\cdot,i^*)}$ and $\text{logit}\left(\phi_{A}^{*(\cdot,i^*)}\right)$ for subdive types $i^* = 1,2,3$. All other parameters and probability transition matrices are assumed to be the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is added to the negative log-likelihood of the maximum likelihood estimate $\left(\hat \mu_{A_x}^{*(\cdot,i^*)}, \hat \sigma_{A_x}^{*(\cdot,i^*)}, \hat \phi_{A}^{*(\cdot,i^*)}\right)$, which is denoted by a red star.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the $y$-acceleration parameters $\mu_{A_y}^{*(\cdot,i^*)}, \sigma_{A_y}^{*(\cdot,i^*)}$ and $\text{logit}\left(\phi_{A}^{*(\cdot,i^*)}\right)$ for subdive types $i^* = 1,2,3$. All other parameters and probability transition matrices are assumed to be the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is added to the negative log-likelihood of the maximum likelihood estimate $\left(\hat \mu_{A_y}^{*(\cdot,i^*)}, \hat \sigma_{A_y}^{*(\cdot,i^*)}, \hat \phi_{A}^{*(\cdot,i^*)}\right)$, which is denoted by a red star.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the $z$-acceleration parameters $\mu_{A_z}^{*(\cdot,i^*)}, \sigma_{A_z}^{*(\cdot,i^*)}$ and $\text{logit}\left(\phi_{A}^{*(\cdot,i^*)}\right)$ for subdive types $i^* = 1,2,3$. All other parameters and probability transition matrices are assumed to be the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is added to the negative log-likelihood of the maximum likelihood estimate $\left(\hat \mu_{A_z}^{*(\cdot,i^*)}, \hat \sigma_{A_z}^{*(\cdot,i^*)}, \hat \phi_{A}^{*(\cdot,i^*)}\right)$, which is denoted by a red star.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the fine-scale probability transition matrix parameters associated both dive types ($\eta^{*(i)}_{i^*,j^*}$ for dive types $i = 1,2$ and subdive states $i^*,j^* = 1,2,3$). Recall that $\eta^{*(i)}_{i^*,i^*} = 0$ and that $\Gamma^{*(i)}_{i^*,j^*} = \exp(\eta^{*(i)}_{i^*,j^*}) / \sum_{k^*} \exp(\eta^{*(i)}_{i^*,k^*})$. All other parameters and probability transition matrices are assumed to be the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is added to the negative log-likelihood of the maximum likelihood estimate $\hat \Gamma^{*(i)}$, which is denoted by a red star. Note that the long ridges of approximately equal likelihood are due to the fact that $\hat \Gamma^{*(i)}_{1,3} \approx \hat \Gamma^{*(i)}_{3,1} \approx 0$. This indicates that the whale rarely switches directly between subdive states 1 and 3.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: The $x$--component of acceleration $\left(y^*_{t,t^*}\right)_x$ (top two panels) and dive depth (bottom two panels) of a northern resident killer whale for a sequence of six selected dives. Each panel is partitioned into dives by vertical black lines. The curve colours in the first and third panels correspond to the estimated dive types while the curve colours of the second and fourth panels correspond the estimated subdive states. Both the dive types and subdive states are estimated by fitting the CarHHMM-DFT to the data and performing the forward-backward algorithm to determine the hidden state with the highest probability.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: The $x$--component of acceleration $\left(y^*_{t,t^*}\right)_x$ (top two panels) and dive depth (bottom two panels) of a northern resident killer whale for a sequence of six selected dives. Each panel is partitioned into dives by vertical black lines. The curve colours in the first and third panels correspond to the estimated dive types while the curve colours of the second and fourth panels correspond the estimated subdive states. Both the dive types and subdive states are estimated by fitting the HHMM-DFT to the data and performing the forward-backward algorithm to determine the hidden state with the highest probability.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: The $x$--component of acceleration $\left(y^*_{t,t^*}\right)_x$ (top two panels) and dive depth (bottom two panels) of a northern resident killer whale for a sequence of six selected dives. Each panel is partitioned into dives by vertical black lines. The curve colours in the first and third panels correspond to the estimated dive types while the curve colours of the second and fourth panels correspond the estimated subdive states. Both the dive types and subdive states are estimated by fitting the CarHHMM to the data and performing the forward-backward algorithm to determine the hidden state with the highest probability.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: The $x$--component of acceleration $\left(y^*_{t,t^*}\right)_x$ (top two panels) and dive depth (bottom two panels) of a northern resident killer whale for a sequence of six selected dives. Each panel is partitioned into dives by vertical black lines. The curve colours in the first and third panels correspond to the estimated dive types while the curve colours of the second and fourth panels correspond the estimated subdive states. Both the dive types and subdive states are estimated by fitting the CarHMM-DFT to the data and performing the forward-backward algorithm to determine the hidden state with the highest probability.
        \addtocounter{fignum}{1}

>>>>

    \section{Likelihood Surface of the CarHHMM-DFT}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the dive duration parameters $\mu_Y^{(i)}$ and $\sigma_Y^{(i)}$ for dive types $i = 1,2$. All other parameters and probability transition matrices are set equal to the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is added to the negative log-likelihood of the maximum likelihood estimate $(\hat \mu_Y,\hat \sigma_Y)$, which is denoted by a red star.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the coarse-scale probability transition matrix parameters $\eta_{1,2}$ and $\eta_{2,1}$. Recall that $\Gamma_{i,j} = \exp(\eta_{i,j}) / \sum_k \exp(\eta_{i,k})$. All other parameters and probability transition matrices are set equal to the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is added to the negative log-likelihood of the maximum likelihood estimate $(\hat \eta_{1,2},\hat \eta_{2,1})$, which is denoted by a red star.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the wiggliness parameters $\mu_W^{*(\cdot,i^*)}$ and $\sigma_W^{*(\cdot,i^*)}$ for subdive types $i^* = 1,2,3$. All other parameters and probability transition matrices are set equal to the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is added to the negative log-likelihood of the maximum likelihood estimate $(\hat \mu_W^{*(\cdot,i^*)}, \hat \sigma_W^{*(\cdot,i^*)})$, which is denoted by a red star.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the $x$-acceleration parameters $\mu_{A_x}^{*(\cdot,i^*)}, \sigma_{A_x}^{*(\cdot,i^*)}$ and logit of $\phi_{A}^{*(\cdot,i^*)}$ for subdive types $i^* = 1,2,3$. All other parameters and probability transition matrices are set equal to the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is added to the negative log-likelihood of the maximum likelihood estimate $(\hat \mu_{A_x}^{*(\cdot,i^*)}, \hat \sigma_{A_x}^{*(\cdot,i^*)}, \hat \phi_{A}^{*(\cdot,i^*)})$, which is denoted by a red star.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the $y$-acceleration parameters $\mu_{A_y}^{*(\cdot,i^*)}, \sigma_{A_y}^{*(\cdot,i^*)}$ and logit of $\phi_{A}^{*(\cdot,i^*)}$ for subdive types $i^* = 1,2,3$. All other parameters and probability transition matrices are set equal to the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is added to the negative log-likelihood of the maximum likelihood estimate $(\hat \mu_{A_y}^{*(\cdot,i^*)}, \hat \sigma_{A_y}^{*(\cdot,i^*)}, \hat \phi_{A}^{*(\cdot,i^*)})$, which is denoted by a red star.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the $z$-acceleration parameters $\mu_{A_z}^{*(\cdot,i^*)}, \sigma_{A_z}^{*(\cdot,i^*)}$ and logit of $\phi_{A}^{*(\cdot,i^*)}$ for subdive types $i^* = 1,2,3$. All other parameters and probability transition matrices are set equal to the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is added to the negative log-likelihood of the maximum likelihood estimate $(\hat \mu_{A_z}^{*(\cdot,i^*)}, \hat \sigma_{A_z}^{*(\cdot,i^*)}, \hat \phi_{A}^{*(\cdot,i^*)})$, which is denoted by a red star.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the fine-scale probability transition matrix parameters associated both dive types ($\eta^{*(i)}_{i^*,j^*}$ for dive types $i = 1,2$ and subdive states $i^*,j^* = 1,2,3$). Recall that $\eta^{*(i)}_{i^*,i^*} = 0$ and that $\Gamma^{*(i)}_{i^*,j^*} = \exp(\eta^{*(i)}_{i^*,j^*}) / \sum_{k^*} \exp(\eta^{*(i)}_{i^*,k^*})$. All other parameters and probability transition matrices are set equal to the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is added to the negative log-likelihood of the maximum likelihood estimate $\hat \Gamma^{*(i)}$, which is denoted by a red star. Note that the long ridges of approximately equal likelihood are due to the fact that $\hat \Gamma^{*(i)}_{1,3} \approx \hat \Gamma^{*(i)}_{3,1} \approx 0$. This indicates that the whale rarely switches directly between subdive states 1 and 3.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: The $x$--component of acceleration $(y^*_{t,t^*})_x$ (top two panels) and dive depth (bottom two panels) of a northern resident killer whale for a sequence of six selected dives. Each panel is partitioned into dives by vertical black lines. The curve colours in the first and third panels correspond to the estimated dive types while the curve colours of the second and fourth panels correspond the estimated subdive states. Both the dive types and subdive states are estimated by fitting the CarHHMM-DFT to the data and performing the forward-backward algorithm to determine the hidden state with the highest probability.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: The $x$--component of acceleration $(y^*_{t,t^*})_x$ (top two panels) and dive depth (bottom two panels) of a northern resident killer whale for a sequence of six selected dives. Each panel is partitioned into dives by vertical black lines. The curve colours in the first and third panels correspond to the estimated dive types while the curve colours of the second and fourth panels correspond the estimated subdive states. Both the dive types and subdive states are estimated by fitting the HHMM-DFT to the data and performing the forward-backward algorithm to determine the hidden state with the highest probability.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: The $x$--component of acceleration $(y^*_{t,t^*})_x$ (top two panels) and dive depth (bottom two panels) of a northern resident killer whale for a sequence of six selected dives. Each panel is partitioned into dives by vertical black lines. The curve colours in the first and third panels correspond to the estimated dive types while the curve colours of the second and fourth panels correspond the estimated subdive states. Both the dive types and subdive states are estimated by fitting the CarHHMM to the data and performing the forward-backward algorithm to determine the hidden state with the highest probability.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: The $x$--component of acceleration $(y^*_{t,t^*})_x$ (top two panels) and dive depth (bottom two panels) of a northern resident killer whale for a sequence of six selected dives. Each panel is partitioned into dives by vertical black lines. The curve colours in the first and third panels correspond to the estimated dive types while the curve colours of the second and fourth panels correspond the estimated subdive states. Both the dive types and subdive states are estimated by fitting the CarHMM-DFT to the data and performing the forward-backward algorithm to determine the hidden state with the highest probability.
        \addtocounter{fignum}{1}
